\malg: a novel interest management algorithm for distributed simulations of MMOGs

### 2009-07-08 trunk/textos/sbgames2009/aoisbgames2008.tex

\documentclass[conference]{IEEEtran}

@@

%\DeclareGraphicsExtensions{eps}

%\newcommand{\malg}{A$^3$}
\newcommand{\malg}{A$^3$}
\newcommand{\cridis}{critical\_distance}
\newcommand{\viewdis}{view\_distance}
\newcommand{\circleaoi}{C}
\newcommand{\circleat}{C \& A}
\newcommand{\fov}{FoV}
\newcommand{\noaoi}{None}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{\malg: a novel interest management algorithm for distributed simulations of MMOGs}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Carlos Eduardo Benevides Bezerra}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: carlos.bezerra@inf.ufrgs.br}
% \and
% \IEEEauthorblockN{Fábio Reis Cecin}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: frcecin@inf.ufrgs.br}
% \and
% \IEEEauthorblockN{Cláudio Fernando Resin Geyer}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: geyer@inf.ufrgs.br}
% }

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%AQUI OS AUTORES
\author{\IEEEauthorblockN{Carlos Eduardo B. Bezerra,
	Fábio R. Cecin,
	Cláudio F. R. Geyer}
\IEEEauthorblockA{Federal University of Rio Grande do Sul\\
	Bento Gonçalves, 9500, Porto Alegre, RS, Brazil,\\
	Email: \{carlos.bezerra, fcecin, geyer\}@inf.ufrgs.br}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath

% :: OLD ABSTRACT ::
Traditionally, a central server is utilized to provide support to MMOGs (massively multiplayer online games), where the number of participants is in the order of tens of thousands. Much work has been done trying to create a fully peer-to-peer model to support this kind of application, in order to minimize the maintenance cost of its infrastructure, but critical questions remain. Examples of the problems relative to peer-to-peer MMOG support systems are: vulnerability to cheating, overload of the upload links of the peers and difficulty to maintain consistency of the simulation among the participants. In this work, we propose the utilization of geographically distributed lower-cost nodes, working as a distributed game server. The distribution model and some related works are also presented. To address the communication cost imposed to the servers, we specify the \malg\ algorithm, which is a novel refinement of the interest management technique, significantly reducing the necessary bandwidth. Simulations have been made with ns-2 and their results demonstrate that our approach achieves the least bandwidth utilization, with a 33.10\% maximum traffic reduction and 33.58\% average traffic reduction, when compared to other area of interest algorithms.

% :: ABSTRACT++ BY FABIO ::
%Traditionally, a centralized server or cluster of server machines is employed to provide support for MMOGs (massively multiplayer online games), where the number of participants is in the order of tens of thousands. In order to minimize the maintenance cost of this infrastructure, much work has been done trying to create a fully decentralized or peer-to-peer model to support MMOGs, but some critical issues remain unsolved. Since there is no general availability of Internet-wide IP multicast routing, decentralization proposals usually require peers on home broadband links to upload real-time game traffic to several peers simultaneously, and widespread last-mile broadband technologies such as ADSL have low upstream capacities by design which are usually further capped by connectivity providers. This adds to the difficulty of maintaining the game simulation consistent among the peers, no matter how they are to be organized. In this work, we propose the use of geographically distributed lower-cost nodes, served by higher-end but still consumer-grade broadband links, which work together to act as a distributed MMOG serving entity. The distribution model and some related works are also presented. To address the communication cost imposed on those `super-peer' server nodes, we also propose a novel refinement to the area of interest technique which significantly reduces the bandwidth consumed by game update traffic. Simulations have been made with ns-2, comparing different area of interest algorithms. The results show that our approach achieves the least bandwidth utilization, with a 33.10\% maximum traffic reduction and 33.58\% average traffic reduction, when compared to other area of interest algorithms.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
In the last years, electronic games have become very popular, specially MMOGs (massively multiplayer online games), where the number of simultaneous players is in the order tens of thousands \cite{cecin2004fsa}. World of Warcraft \cite{worldofwarcraft}, Lineage II \cite{lineage2} and Guild Wars \cite{guildwars} are successful examples of MMOGs.

@@

In the past few years, many works have been done trying to distribute the network support for massively multiplayer online games. One approach is the peer-to-peer model \cite{crippa2007pps, hampel2006ppa, knutsson2004pps}, which has some issues concerning game state consistency among the participating peers, vulnerability to cheating, and increased bandwidth usage by each peer. There are some proposals that try to minimize these problems. One of them is presented in \cite{schiele2007rpp}, where it is suggested the division of the virtual simulated game environment into regions, and for each region, a peer is chosen to be its coordinator. The function of this peer will be manage the players' interest. It will check who really needs each state update, so each player does not receive every update packet. This way, the peers' bandwidth usage is reduced. However, it will be still higher than with a server performing all the broadcasts. In the peer-to-peer model, each player usually must send updates to more than one other player. Besides, the chosen peer must be trustable to be the region coordinator.

@@

A key problem in the peer-to-peer architectures, in what concerns the use of interest management, is that some peers are responsible for part of the simulation and for deciding to whom each state update interests. Assuming that there are only trustable players, the interest management technique may be useful. However, supposing that a player is malicious, he could hack his client software and avoid sending state updates to certain players, who would be in disadvantage. Consequently, the distributed server model \cite{assiotis2006dam, ng2002msa, fiedler2002cam, rieche2007ppb} is considered more suitable to use the interest management technique.

@@

Regarding the interest management technique, works such as \cite{morse1996iml, rak1996egb, zou2001egt, morgan2005imm} and \cite{minson2005aim} may be cited. In \cite{rak1996egb}, it is proposed an interest management scheme based on a virtual environment divided into a grid of cells. Each cell is assigned to a multicast group. Each participant of the simulation subscribes then to the multicast group of the cell where he is, as well as of the neighboring cells if they are within his field of view. Each participant then sends his state updates to the multicast group of the region where he is situated.

@@

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

@@

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{dsrt}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \begin{thebibliography}{1}
% 
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
% 
% \end{thebibliography}

% that's all folks
\end{document}

>>>>

\documentclass[a4paper]{sbgames}               % final
%\usepackage[scaled=.92]{helvet}
\usepackage{times}
\usepackage{graphicx}

%% use this for zero \parindent and non-zero \parskip, intelligently.
\usepackage{parskip}

%% the 'caption' package provides a nicer-looking replacement
\usepackage[labelfont=bf,textfont=it]{caption}

\usepackage{url}

@@

\newcommand{\malg}{A$^3$}
\newcommand{\cridis}{critical\_distance}
\newcommand{\viewdis}{view\_distance}
\newcommand{\circleaoi}{C}
\newcommand{\circleat}{C \& A}
\newcommand{\fov}{FoV}
\newcommand{\noaoi}{None}

%% Paper title.
\title{\malg: a novel interest management algorithm for distributed simulations of MMOGs}

%% Author and Affiliation (multiple authors). Use: and between authors

\author{Carlos Eduardo B. Bezerra\\ Fábio R. Cecin\\ Cláudio F. R. Geyer\\ Federal University of Rio Grande do Sul
}
\contactinfo{\{carlos.bezerra,fcecin,geyer\}@inf.ufrgs.br
}
%% Keywords that describe your work.
\keywords{Interest management, distributed simulations, MMOG}

%% Start of the paper
% Attention: As you need to insert EPS images in Postscript, 
% you need to insert PDF images into PDFs. 
% In the text, extensions cancbe omitted (latex use .eps, pdflatex get .pdf) 
% To convert them: epstopdf myimage.eps
\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

\maketitle

%% Abstract section.

\begin{abstract}

  Traditionally, a central server is utilized to provide support to MMOGs (massively multiplayer online games), where the number of participants is in the order of tens of thousands. Much work has been done trying to create a fully peer-to-peer model to support this kind of application, in order to minimize the maintenance cost of its infrastructure, but critical questions remain. Examples of the problems relative to peer-to-peer MMOG support systems are: vulnerability to cheating, overload of the upload links of the peers and difficulty to maintain consistency of the simulation among the participants. In this work, we propose the utilization of geographically distributed lower-cost nodes, working as a distributed game server. The distribution model and some related works are also presented. To address the communication cost imposed to the servers, we specify the \malg\ algorithm, which is a novel refinement of the interest management technique, significantly reducing the necessary bandwidth. Simulations have been made with ns-2 and their results demonstrate that our approach achieves the least bandwidth utilization, with a 33.10\% maximum traffic reduction and 33.58\% average traffic reduction, when compared to other area of interest algorithms.
\end{abstract}

%% The ``\keywordlist'' command prints out the keywords.
\keywordlist
\contactlist

In the last years, electronic games have become very popular, specially MMOGs (massively multiplayer online games), where the number of simultaneous players is in the order tens of thousands \cite{cecin2004fsa}. World of Warcraft \cite{worldofwarcraft}, Lineage II \cite{lineage2} and Guild Wars \cite{guildwars} are successful examples of MMOGs.

@@

In the past few years, many works have been done trying to distribute the network support for massively multiplayer online games. One approach is the peer-to-peer model \cite{crippa2007pps} \cite{hampel2006ppa} \cite{knutsson2004pps}, which has some issues concerning game state consistency among the participating peers, vulnerability to cheating, and increased bandwidth usage by each peer. There are some proposals that try to minimize these problems. One of them is presented in \cite{schiele2007rpp}, where it is suggested the division of the virtual simulated game environment into regions, and for each region, a peer is chosen to be its coordinator. The function of this peer will be manage the players' interest. It will check who really needs each state update, so each player does not receive every update packet. This way, the peers' bandwidth usage is reduced. However, it will be still higher than with a server performing all the broadcasts. In the peer-to-peer model, each player usually must send updates to more than one other player. Besides, the chosen peer must be trustable to be the region coordinator.

@@

A key problem in the peer-to-peer architectures, in what concerns the use of interest management, is that some peers are responsible for part of the simulation and for deciding to whom each state update interests. Assuming that there are only trustable players, the interest management technique may be useful. However, supposing that a player is malicious, he could hack his client software and avoid sending state updates to certain players, who would be in disadvantage. Consequently, the distributed server model \cite{assiotis2006dam} \cite{ng2002msa} \cite{fiedler2002cam} \cite{rieche2007ppb} is considered more suitable to use the interest management technique.

@@

Regarding the interest management technique, works such as \cite{morse1996iml}, \cite{rak1996egb}, \cite{zou2001egt}, \cite{morgan2005imm} and \cite{minson2005aim} may be cited. In \cite{rak1996egb}, it is proposed an interest management scheme based on a virtual environment divided into a grid of cells. Each cell is assigned to a multicast group. Each participant of the simulation subscribes then to the multicast group of the cell where he is, as well as of the neighboring cells if they are within his field of view. Each participant then sends his state updates to the multicast group of the region where he is situated.

@@

\section*{Acknowledgment}

@@

\bibliographystyle{sbgames}
\bibliography{aoisbgames2008}
\end{document}
